Looking for a Signal in the Noise: Revisiting Obesity and the Bacterial Microbiome

### 2016-02-24 Sze_ObesityMicrobiomeMeta_PeerJ_2016.Rmd
adding updated Amish data

We used the categorical variable of obese or non-obese to analyze each data set seperately for significant correlations to Bacteroidetes, Firmicutes, Bacteroidetes/Firmicute (B/F) ratio, Shannon diversity, OTU richness, evenness, and the Bray-Curtis distance matrix which was a similar approach taken by two previous reviews [@walters_meta-analyses_2014; @finucane_taxonomic_2014].  Using a P-value cutoff for significance as less than 0.05 we found a total of 1 significant result for Shannon diversity, 3 significant results for OTU richness, and 1 signficant result for the Bray-Curtis distance matrix.  There were no significant results for Bacteroidetes, Firmicutes, or the B/F ratio.  The full summary of the exact P-values for each respective study and variable tested can be found in table 3.

Next we investigated how well the OTUs in each study were able to classify obese and non-obese individuals using the random forest algorithm.  We found that the average Out Of Bag (OOB) Area Under the Curve (AUC) was 0.7553 +/- 0.0888.  We found that the total range for the total number of variables used for the classification ranged from 3 (Ross and Escobar) to 32 (Zupancic).  The total number of trees and node size was standardized to 1000 and 20 respectively for each data set.  One aspect we found surprising was that there was very little overlap in the variables used by the random forest algorithm for the classification of obese and non-obese individuals by study.  The three most common variables was for an OTU that could be classified to the family Ruminococcaceae (7/9 studies), an OTU that could be classified to the family Lachnospiraceae (5/9 studies), and the Shannon diversity metric (4/9 studies).

@@

We next moved on to the pooled meta-analysis.  For the Shannon diversity RR, all studies but Wu et al had those with a lower Shannon diversity then the median as being higher risk for obesity [Figure 2].  This was confirmed with the pooled analysis showing an increased RR of 1.19 (CI 1.04, 1.36) for obesity in the lower then median Shannon diversity group (P-value = 0.0139).  When investigated the B/F ratio RR most studies showed that a low B/F ratio tended to have less obese individuals than those with a high B/F ratio [Figure 3].  A notable exception to this was the HMP data set and the Baxter data set [Figure 3].  The pooled RR was 0.86 (CI 0.75, 0.99) suggesting that a low B/F ratio does indeed have less obese individuals (P-value = 0.0296).

@@

The final forest plot that we used for this analysis investigated how the different OTUs, classified to Ruminococcaceae, that were important for the classification of obese and non-obese based on the OOB AUC in the the seperate study analysis looked when pooled together to assess the RR [Figure S4].  There was no significant increase in RR based on the Ruminococcaceae OTU (RR = 0.98 (CI 0.63, 1.53), P-value = 0.9225).  Thus it would seem that although Ruminococcaceae is important for classification the directionality is study dependent [Figure S4].

@@

Fourth, although a significant result exists for the RR for both Shannon diversity and the B/F ratio, it is small.  This is also the case for the ZScore normalization analysis of the Shannon diversity.  As an example, a lower than median Shannon diversity only has an overall 19% increase of obesity in it versus the higher than median Shannon diversity group.  So even though there is a detectable signal it remains to be seen if this is a biologically significant phenomenon that could have a tangible effect on obesity.

@@

#Read in and match metadata to microbiome data
metadata <- read.csv("amish_obesity_table2.csv")
test <- metadata[!duplicated(metadata$submitted_sample_id_s), ]
rownames(test) <- test[, 4]
shared.data <- read.table("combined.0.03.subsample.shared", header=T)
rownames(shared.data) <- shared.data[, 2]
shared.data <- shared.data[, -c(1:3)]
keep  <- rownames(shared.data)

test2 <- test[keep, ]
good.metadata <- test2
test3 <- test2[!duplicated(test2$submitted_subject_id_s), ]
keep <- rownames(test3)
microbiome <- shared.data[keep, ]
metadata <- test3

rm(good.metadata, shared.data, keep, test, test2, test3)

@@

#Create Obese Yes/No groups
# One problem is that we are assuming those with "not provided" are defaulting to the not affected group
metadata$obese[metadata$subject_is_affected_s=="<not provided>" | metadata$subject_is_affected_s=="No"] <- "No"
metadata$obese[metadata$subject_is_affected_s=="Yes"] <- "Yes"

#create groups to be used
obese <- factor(metadata$obese)

@@

zupancicH <- anova(lm(H ~ obese)) #P-value=0.2034
zupancicS <- anova(lm(S ~ obese)) #P-value=0.03516
zupancicJ <- anova(lm(J ~ obese)) #P-value=0.5369

@@

zupancicBacter <- anova(lm(bacter ~ obese)) #P-value=0.9582
zupancicFirm <- anova(lm(firm ~ obese)) #P-value=0.2654
zupancicBF <- anova(lm(BFratio ~ obese)) #P-value=0.9899

@@

set.seed(3)
zupancic2 <- adonis(microbiome ~ obese, permutations=1000) 
zupancicPERM <- zupancic2$aov.tab
#PERMANOVA=0.5734, pseudo-F=0.95626

@@

zupancicHEpi <- epi.2by2(r.test, method="cohort.count")
zupancicHMassoc <- zupancicHEpi$massoc
zupancicHRR <- zupancicHMassoc$RR.strata.score
zupancicHRRsig <- zupancicHMassoc$chisq.strata
## Risk Ratio = 1.50
## CI = 1.04, 2.16
## p-value = 0.025

@@

zupancicBFEpi <- epi.2by2(r.test, method="cohort.count")
zupancicBFMassoc <- zupancicBFEpi$massoc
zupancicBFRR <- zupancicBFMassoc$RR.strata.score
zupancicBFRRsig <- zupancicBFMassoc$chisq.strata
## Risk Ratio = 0.89
## CI = 0.63, 1.27
## p-value = 0.53

@@

#Create Obese.num group
obese <- factor(metadata$obese.num)

@@

#Try AUCRF with default measures provided in readme
#set.seed(3)
#zupancicAUCFit <- AUCRF(obese ~ ., data=testset, ntree=1000, nodesize=20)
# list of 42 Measures, AUCopt = 0.6365534

@@

StudyNo <- c(1, 2, 3, 4, 5, 6, 7, 8, 9)
Study <- c("Baxter", "HMP", "Ross", "Escobar", "Arumugam", "Turnbaugh", "Zupancic", "Wu", "Goodrich")
Year <- c("2015", "2011", "2015", "2014", "2010", "2009", "2012", "2012", "2014")
Total <- c("n=172", "n=256", "n=63", "n=30", "n=85", "n=146", "n=207", "n=58", "n=507")
tpos <- c(BaxLowShannonGroup[2, 2], HMPLowShannonGroup[2, 2], RossLowShannonGroup[2, 2], EscoLowShannonGroup[2, 2], AruLowShannonGroup[2, 2], TurnLowShannonGroup[2, 2], ZupaLowShannonGroup[2, 2], WuLowShannonGroup[2, 2], GoodLowShannonGroup[2, 2]) #obese in low diversity group
tneg <- c(BaxLowShannonGroup[1, 2], HMPLowShannonGroup[1, 2], RossLowShannonGroup[1, 2], EscoLowShannonGroup[1, 2], AruLowShannonGroup[1, 2], TurnLowShannonGroup[1, 2], ZupaLowShannonGroup[1, 2], WuLowShannonGroup[1, 2], GoodLowShannonGroup[1, 2]) #nonobese in low diversity group
cpos <- c(BaxHighShannonGroup[2, 2], HMPHighShannonGroup[2, 2], RossHighShannonGroup[2, 2], EscoHighShannonGroup[2, 2], AruHighShannonGroup[2, 2], TurnHighShannonGroup[2, 2], ZupaHighShannonGroup[2, 2], WuHighShannonGroup[2, 2], GoodHighShannonGroup[2, 2]) #obese in high diversity group
cneg <- c(BaxHighShannonGroup[1, 2], HMPHighShannonGroup[1, 2], RossHighShannonGroup[1, 2], EscoHighShannonGroup[1, 2], AruHighShannonGroup[1, 2], TurnHighShannonGroup[1, 2], ZupaHighShannonGroup[1, 2], WuHighShannonGroup[1, 2], GoodHighShannonGroup[1, 2]) #nonobese in high diversity group
RR <- signif(c(baxterHRR[1,1], HMPHRR[1,1], rossHRR[1,1], escobarHRR[1,1], ArumugamHRR[1,1], turnbaughHRR[1,1], zupancicHRR[1,1], WuHRR[1,1], goodrichHRR[1,1]), 2)
low <- signif(c(baxterHRR[1,2], HMPHRR[1,2], rossHRR[1,2], escobarHRR[1,2], ArumugamHRR[1,2], turnbaughHRR[1,2], zupancicHRR[1,2], WuHRR[1,2], goodrichHRR[1,2]), 2)
high <- signif(c(baxterHRR[1,3], HMPHRR[1,3], rossHRR[1,3], escobarHRR[1,3], ArumugamHRR[1,3], turnbaughHRR[1,3], zupancicHRR[1,3], WuHRR[1,3], goodrichHRR[1,3]), 2)

@@

**Figure 2: Meta Analysis of the Relative Risk of Obesity Based on Shannon Diversity.**  Groups were divided for each study on high and low Shannon Diversity groups based on the median for that study.  The overall pooled relative risk was 1.19 for the low diversity group (P-value = 0.0139).

@@

StudyNo <- c(1, 2, 3, 4, 5, 6, 7, 8, 9)
Study <- c("Baxter", "HMP", "Ross", "Escobar", "Arumugam", "Turnbaugh", "Zupancic", "Wu", "Goodrich")
Year <- c("2015", "2011", "2015", "2014", "2010", "2009", "2012", "2012", "2014")
Total <- c("n=172", "n=256", "n=63", "n=30", "n=85", "n=146", "n=201", "n=58", "n=507")
tpos <- c(BaxLowBFGroup[2, 2], HMPLowBFGroup[2, 2], RossLowBFGroup[2, 2], EscoLowBFGroup[2, 2], AruLowBFGroup[2, 2], TurnLowBFGroup[2, 2], ZupaLowBFGroup[2, 2], WuLowBFGroup[2, 2], GoodLowBFGroup[2, 2]) #obese in low BF group
tneg <- c(BaxLowBFGroup[1, 2], HMPLowBFGroup[1, 2], RossLowBFGroup[1, 2], EscoLowBFGroup[1, 2], AruLowBFGroup[1, 2], TurnLowBFGroup[1, 2], ZupaLowBFGroup[1, 2], WuLowBFGroup[1, 2], GoodLowBFGroup[1, 2]) #nonobese in low BF group
cpos <- c(BaxHighBFGroup[2, 2], HMPHighBFGroup[2, 2], RossHighBFGroup[2, 2], EscoHighBFGroup[2, 2], AruHighBFGroup[2, 2], TurnHighBFGroup[2, 2], ZupaHighBFGroup[2, 2], WuHighBFGroup[2, 2], GoodHighBFGroup[2, 2]) #obese in high BF group
cneg <- c(BaxHighBFGroup[1, 2], HMPHighBFGroup[1, 2], RossHighBFGroup[1, 2], EscoHighBFGroup[1, 2], AruHighBFGroup[1, 2], TurnHighBFGroup[1, 2], ZupaHighBFGroup[1, 2], WuHighBFGroup[1, 2], GoodHighBFGroup[1, 2]) #nonobese in high BF group
RR <- signif(c(baxterBFRR[1,1], HMPBFRR[1,1], rossBFRR[1,1], escobarBFRR[1,1], ArumugamBFRR[1,1], turnbaughBFRR[1,1], zupancicBFRR[1,1], WuBFRR[1,1], goodrichBFRR[1,1]), 2)
low <- signif(c(baxterBFRR[1,2], HMPBFRR[1,2], rossBFRR[1,2], escobarBFRR[1,2], ArumugamBFRR[1,2], turnbaughBFRR[1,2], zupancicBFRR[1,2], WuBFRR[1,2], goodrichBFRR[1,2]), 2)
high <- signif(c(baxterBFRR[1,3], HMPBFRR[1,3], rossBFRR[1,3], escobarBFRR[1,3], ArumugamBFRR[1,3], turnbaughBFRR[1,3], zupancicBFRR[1,3], WuBFRR[1,3], goodrichBFRR[1,3]), 2)

@@

**Figure 3: Meta Analysis of the Relative Risk of Obesity Based on Bacteroidetes/Firmicutes Ratio.**  Groups were divided for each study on high and low B/F ratio groups based on the median for that specific study.  The overall pooled relative risk was 0.86 for the low diversity group (P-value = 0.0296).

@@

#Load in BMI data
data1 <- read.csv("combined.H.corr.data.csv")
#Load in Turnbaugh data
turn.data <- read.csv("Turnbaugh.normalized.csv")
#Load in Amish data
amish.data <- read.csv("Amish.normalized.csv")

#Amish STD = 0.5893173
amish.ZScore.H <- amish.data$H.corr/0.5893173
#Turnbaugh STD = 0.4818751
turn.ZScore.H <- turn.data$H.corr/0.4818751
#MetaHit-Den STD = 0.4038558
keepmetahit <- which(data1$dataset == "MetaHit")
metahitdata <- data1[keepmetahit, ]
metahit.ZScore.H <- metahitdata$H.corr/0.4038558
#Korean STD = 0.5400615
keepKorean <- which(data1$dataset == "Korean")
Koreandata <- data1[keepKorean, ]
Korean.ZScore.H <- Koreandata$H.corr/0.5400615
#HMP STD = 0.6774514
keepHMP <- which(data1$dataset == "HMP")
hmpdata <- data1[keepHMP, ]
HMP.ZScore.H <- hmpdata$H.corr/0.6774514
#CCHC STD = 0.5784522
keepCCHC <- which(data1$dataset == "CCHC")
CCHCdata <- data1[keepCCHC, ]
CCHC.ZScore.H <- CCHCdata$H.corr/0.5784522
#GLNE07 STD = 0.4663871
keepGLNE <- which(data1$dataset == "GLNE")
GLNEdata <- data1[keepGLNE, ]
GLNE.ZScore.H <- GLNEdata$H.corr/0.4663871
#Escobar STD = 0.5378935
keepescobar <- which(data1$dataset == "Escobar")
escobardata <- data1[keepescobar, ]
Escobar.ZScore.H <- escobardata$H.corr/0.5378935
#COMBO STD = 0.6870763
keepCOMBO <- which(data1$dataset == "COMBO")
COMBOdata <- data1[keepCOMBO, ]
COMBO.ZScore.H <- COMBOdata$H.corr/0.6870763
#COMBO STD = 0.3489582
keeptwinsUK <- which(data1$dataset == "twinsUK")
twinsUKdata <- data1[keeptwinsUK, ]
twinsUK.ZScore.H <- twinsUKdata$H.corr/0.3489582

#data1 order Escobar, GLNE, CCHC, HMP, Korean, MetaHit, COMBO, twinsUK
HZScore <- c(Escobar.ZScore.H, GLNE.ZScore.H, CCHC.ZScore.H, HMP.ZScore.H, Korean.ZScore.H, metahit.ZScore.H, COMBO.ZScore.H, twinsUK.ZScore.H)
data1 <- cbind(HZScore, data1)

amish.data <- cbind(amish.ZScore.H, amish.data)
turn.data <- cbind(turn.ZScore.H, turn.data)
rm(Escobar.ZScore.H, GLNE.ZScore.H, CCHC.ZScore.H, HMP.ZScore.H, Korean.ZScore.H, metahit.ZScore.H, amish.ZScore.H, turn.ZScore.H, COMBO.ZScore.H, twinsUK.ZScore.H)
rm(keepmetahit, keepCCHC, keepescobar, keepGLNE, keepHMP, keepKorean)

@@

#Create group analysis table to harness full data set
#For ZScore normalization
d1.obese <- data1$obese
t.obese <- as.character(turn.data$obese)
am.obese <- as.character(amish.data$obese)

@@

d1.H.Zscore <- data1$HZScore
t.H.Zscore <- turn.data$turn.ZScore.H
am.H.Zscore<- amish.data$amish.ZScore.H

#Combine everything together
comb1.obese <- c(d1.obese, t.obese)
comb2.obese <- c(d1.obese, t.obese, am.obese)
comb1.bmi.class <- c(d1.bmi.class, t.bmi.class)
comb1.H.Zscore <- as.data.frame(c(d1.H.Zscore, t.H.Zscore))
comb2.H.Zscore <- as.data.frame(c(d1.H.Zscore, t.H.Zscore, am.H.Zscore))

group.data.test1 <- cbind(comb1.H.Zscore, comb1.obese, comb1.bmi.class)
group.data.test2 <- cbind(comb2.H.Zscore, comb2.obese)
colnames(group.data.test1) <- c("H.Zscore", "obese", "BMI.class")
colnames(group.data.test2) <- c("H.Zscore", "obese")

@@

##Run Zscore normalized Bacter and Firm overall analysis
#data1 order Escobar, GLNE, CCHC, HMP, Korean, MetaHit
d1.zbf.corr <- data2$ZBF
t.zbf.corr <- turn.data$ZBF
am.zbf.corr <- amish.data$ZBF

@@

comb2.zbf.corr <- as.data.frame(c(d1.zbf.corr, t.zbf.corr, am.zbf.corr))
group.data.test2 <- cbind(comb2.zbf.corr, group.data.test2)
colnames(group.data.test2)[1] <- "zbf"

@@

```{r Figure_5, echo=FALSE, message=FALSE, warning=FALSE}
# Section A
p1 <- ggplot(group.data.test2, aes(obese, H.Zscore)) 
p2 <- p1 + geom_boxplot(aes(fill = obese)) + scale_fill_brewer(palette="Set1")
p3 <- p2 + theme(axis.text=element_text(size=8), axis.title=element_text(size=10, face="bold"))
p4 <- p3 + xlab("Obese Category") + ylab("ZScore Normalized H") 
p5 <- p4 + guides(fill=guide_legend(title="Obese")) + theme(legend.title = element_text(size=8, face="bold"))
p6 <- p5 + theme(axis.line=element_line(colour="black"), 
                 panel.grid.major = element_blank(), 
                 panel.grid.minor = element_blank(), 
                 panel.border = element_blank(), 
                 panel.background = element_blank())
p7 <- p6 + ggtitle('A') + theme(plot.title=element_text(hjust=0, face="bold", size=20))

>>>>

We used the categorical variable of obese or non-obese to analyze each data set seperately for significant correlations to Bacteroidetes, Firmicutes, Bacteroidetes/Firmicute (B/F) ratio, Shannon diversity, OTU richness, evenness, and the Bray-Curtis distance matrix which was a similar approach taken by two previous reviews [@walters_meta-analyses_2014; @finucane_taxonomic_2014].  Using a P-value cutoff for significance as less than 0.05 we found a total of 1 significant result for Shannon diversity, 2 significant results for OTU richness, and 1 signficant result for the Bray-Curtis distance matrix.  There were no significant results for Bacteroidetes, Firmicutes, or the B/F ratio.  The full summary of the exact P-values for each respective study and variable tested can be found in table 3.

Next we investigated how well the OTUs in each study were able to classify obese and non-obese individuals using the random forest algorithm.  We found that the average Out Of Bag (OOB) Area Under the Curve (AUC) was 0.7553 +/- 0.0855.  We found that the total range for the total number of variables used for the classification ranged from 3 (Ross and Escobar) to 40 (Zupancic).  The total number of trees and node size was standardized to 1000 and 20 respectively for each data set.  One aspect we found surprising was that there was very little overlap in the variables used by the random forest algorithm for the classification of obese and non-obese individuals by study.  The three most common variables was for an OTU that could be classified to the family Ruminococcaceae (7/9 studies), an OTU that could be classified to the family Lachnospiraceae (5/9 studies), and the Shannon diversity metric (4/9 studies).

@@

We next moved on to the pooled meta-analysis.  For the Shannon diversity RR, all studies but Wu et al had those with a lower Shannon diversity then the median as being higher risk for obesity [Figure 2].  This was confirmed with the pooled analysis showing an increased RR of 1.18 (CI 1.03, 1.36) for obesity in the lower then median Shannon diversity group (P-value = 0.0187).  When investigated the B/F ratio RR most studies showed that a low B/F ratio tended to have less obese individuals than those with a high B/F ratio [Figure 3].  A notable exception to this was the HMP data set and the Baxter data set [Figure 3].  The pooled RR was 0.86 (CI 0.75, 0.99) suggesting that a low B/F ratio does indeed have less obese individuals (P-value = 0.0376).

@@

The final forest plot that we used for this analysis investigated how the different OTUs, classified to Ruminococcaceae, that were important for the classification of obese and non-obese based on the OOB AUC in the the seperate study analysis looked when pooled together to assess the RR [Figure S4].  There was no significant increase in RR based on the Ruminococcaceae OTU (RR = 0.96 (CI 0.60, 1.52), P-value = 0.8575).  Thus it would seem that although Ruminococcaceae is important for classification the directionality is study dependent [Figure S4].

@@

Fourth, although a significant result exists for the RR for both Shannon diversity and the B/F ratio, it is small.  This is also the case for the ZScore normalization analysis of the Shannon diversity.  As an example, a lower than median Shannon diversity only has an overall 18% increase of obesity in it versus the higher than median Shannon diversity group.  So even though there is a detectable signal it remains to be seen if this is a biologically significant phenomenon that could have a tangible effect on obesity.

@@

#Read in and match metadata to microbiome data
metadata <- read.csv("amish_obesity_table2.csv")
metadata2 <- read.csv("amish.metadata.csv")
test <- metadata[!duplicated(metadata$submitted_sample_id_s), ]
rownames(test) <- test[, 4]
test2 <- metadata2[!duplicated(metadata2$SUBJID), ]
rownames(test2) <- test2[, 1]
shared.data <- read.table("combined.0.03.subsample.shared", header=T)
rownames(shared.data) <- shared.data[, 2]
shared.data <- shared.data[, -c(1:3)]
BactFamily <- read.csv("BacterialFamilyData.csv")
rownames(BactFamily) <- BactFamily[, 1]
BactFamily <- BactFamily[, -1]
keep  <- rownames(shared.data)

test3 <- test[keep, ]
good.metadata <- test3
test4 <- test2[keep, ]
good.metadata2 <- test4

family <- BactFamily[keep, ]
microbiome <- shared.data[keep, ]
metadata <- cbind(good.metadata, good.metadata2)

rm(good.metadata, shared.data, keep, test, test2, test3, test4, BactFamily, good.metadata2, metadata2)

test <- metadata[!duplicated(metadata$submitted_subject_id_s), ]
keep  <- rownames(test)
metadata <- test
family <- family[keep, ]
microbiome <- microbiome[keep, ]

rm(test, keep)

metadata <- metadata[complete.cases(metadata), ]
keep <- rownames(metadata)
family <- family[keep, ]
microbiome <- microbiome[keep, ]

rm(keep)

@@

#Create BMI class 2 groups
metadata$BMI.class2[metadata$BMI<=24] <- "Normal"
metadata$BMI.class2[metadata$BMI>24 & metadata$BMI<30] <- "Overweight"
metadata$BMI.class2[metadata$BMI>=30] <- "Obese"

#create groups to be used
bmi <- metadata$BMI
obese <- factor(metadata$obese)

@@

zupancicH <- anova(lm(H ~ obese)) #P-value=0.2671
zupancicS <- anova(lm(S ~ obese)) #P-value=0.05423
zupancicJ <- anova(lm(J ~ obese)) #P-value=0.6229

@@

zupancicBacter <- anova(lm(bacter ~ obese)) #P-value=0.8592
zupancicFirm <- anova(lm(firm ~ obese)) #P-value=0.2564
zupancicBF <- anova(lm(BFratio ~ obese)) #P-value=0.989

@@

set.seed(3)
zupancic2 <- adonis(microbiome ~ obese, permutations=1000) 
zupancicPERM <- zupancic2$aov.tab
#PERMANOVA=0.6863, pseudo-F=0.92888

@@

zupancicHEpi <- epi.2by2(r.test, method="cohort.count")
zupancicHMassoc <- zupancicHEpi$massoc
zupancicHRR <- zupancicHMassoc$RR.strata.score
zupancicHRRsig <- zupancicHMassoc$chisq.strata
## Risk Ratio = 1.49
## CI = 1.01, 2.19
## p-value = 0.043

@@

zupancicBFEpi <- epi.2by2(r.test, method="cohort.count")
zupancicBFMassoc <- zupancicBFEpi$massoc
zupancicBFRR <- zupancicBFMassoc$RR.strata.score
zupancicBFRRsig <- zupancicBFMassoc$chisq.strata
## Risk Ratio = 0.94
## CI = 0.64, 1.36
## p-value = 0.727

@@

@@

#Try AUCRF with default measures provided in readme
#set.seed(3)
#zupancicAUCFit <- AUCRF(obese ~ ., data=testset, ntree=1000, nodesize=20)
#zupAUC <- fit$`OOB-AUCopt`
# list of 40 Measures, AUCopt = 0.6596899

@@

StudyNo <- c(1, 2, 3, 4, 5, 6, 7, 8, 9)
Study <- c("Baxter", "HMP", "Ross", "Escobar", "Arumugam", "Turnbaugh", "Zupancic", "Wu", "Goodrich")
Year <- c("2015", "2011", "2015", "2014", "2010", "2009", "2012", "2012", "2014")
Total <- c("n=172", "n=256", "n=63", "n=30", "n=85", "n=146", "n=199", "n=58", "n=507")
tpos <- c(BaxLowShannonGroup[2, 2], HMPLowShannonGroup[2, 2], RossLowShannonGroup[2, 2], EscoLowShannonGroup[2, 2], AruLowShannonGroup[2, 2], TurnLowShannonGroup[2, 2], ZupaLowShannonGroup[2, 2], WuLowShannonGroup[2, 2], GoodLowShannonGroup[2, 2]) #obese in low diversity group
tneg <- c(BaxLowShannonGroup[1, 2], HMPLowShannonGroup[1, 2], RossLowShannonGroup[1, 2], EscoLowShannonGroup[1, 2], AruLowShannonGroup[1, 2], TurnLowShannonGroup[1, 2], ZupaLowShannonGroup[1, 2], WuLowShannonGroup[1, 2], GoodLowShannonGroup[1, 2]) #nonobese in low diversity group
cpos <- c(BaxHighShannonGroup[2, 2], HMPHighShannonGroup[2, 2], RossHighShannonGroup[2, 2], EscoHighShannonGroup[2, 2], AruHighShannonGroup[2, 2], TurnHighShannonGroup[2, 2], ZupaHighShannonGroup[2, 2], WuHighShannonGroup[2, 2], GoodHighShannonGroup[2, 2]) #obese in high diversity group
cneg <- c(BaxHighShannonGroup[1, 2], HMPHighShannonGroup[1, 2], RossHighShannonGroup[1, 2], EscoHighShannonGroup[1, 2], AruHighShannonGroup[1, 2], TurnHighShannonGroup[1, 2], ZupaHighShannonGroup[1, 2], WuHighShannonGroup[1, 2], GoodHighShannonGroup[1, 2]) #nonobese in high diversity group
RR <- signif(c(baxterHRR[1,1], HMPHRR[1,1], rossHRR[1,1], escobarHRR[1,1], ArumugamHRR[1,1], turnbaughHRR[1,1], zupancicHRR[1,1], WuHRR[1,1], goodrichHRR[1,1]), 2)
low <- signif(c(baxterHRR[1,2], HMPHRR[1,2], rossHRR[1,2], escobarHRR[1,2], ArumugamHRR[1,2], turnbaughHRR[1,2], zupancicHRR[1,2], WuHRR[1,2], goodrichHRR[1,2]), 2)
high <- signif(c(baxterHRR[1,3], HMPHRR[1,3], rossHRR[1,3], escobarHRR[1,3], ArumugamHRR[1,3], turnbaughHRR[1,3], zupancicHRR[1,3], WuHRR[1,3], goodrichHRR[1,3]), 2)

@@

**Figure 2: Meta Analysis of the Relative Risk of Obesity Based on Shannon Diversity.**  Groups were divided for each study on high and low Shannon Diversity groups based on the median for that study.  The overall pooled relative risk was 1.18 for the low diversity group (P-value = 0.0187).

@@

StudyNo <- c(1, 2, 3, 4, 5, 6, 7, 8, 9)
Study <- c("Baxter", "HMP", "Ross", "Escobar", "Arumugam", "Turnbaugh", "Zupancic", "Wu", "Goodrich")
Year <- c("2015", "2011", "2015", "2014", "2010", "2009", "2012", "2012", "2014")
Total <- c("n=172", "n=256", "n=63", "n=30", "n=85", "n=146", "n=199", "n=58", "n=507")
tpos <- c(BaxLowBFGroup[2, 2], HMPLowBFGroup[2, 2], RossLowBFGroup[2, 2], EscoLowBFGroup[2, 2], AruLowBFGroup[2, 2], TurnLowBFGroup[2, 2], ZupaLowBFGroup[2, 2], WuLowBFGroup[2, 2], GoodLowBFGroup[2, 2]) #obese in low BF group
tneg <- c(BaxLowBFGroup[1, 2], HMPLowBFGroup[1, 2], RossLowBFGroup[1, 2], EscoLowBFGroup[1, 2], AruLowBFGroup[1, 2], TurnLowBFGroup[1, 2], ZupaLowBFGroup[1, 2], WuLowBFGroup[1, 2], GoodLowBFGroup[1, 2]) #nonobese in low BF group
cpos <- c(BaxHighBFGroup[2, 2], HMPHighBFGroup[2, 2], RossHighBFGroup[2, 2], EscoHighBFGroup[2, 2], AruHighBFGroup[2, 2], TurnHighBFGroup[2, 2], ZupaHighBFGroup[2, 2], WuHighBFGroup[2, 2], GoodHighBFGroup[2, 2]) #obese in high BF group
cneg <- c(BaxHighBFGroup[1, 2], HMPHighBFGroup[1, 2], RossHighBFGroup[1, 2], EscoHighBFGroup[1, 2], AruHighBFGroup[1, 2], TurnHighBFGroup[1, 2], ZupaHighBFGroup[1, 2], WuHighBFGroup[1, 2], GoodHighBFGroup[1, 2]) #nonobese in high BF group
RR <- signif(c(baxterBFRR[1,1], HMPBFRR[1,1], rossBFRR[1,1], escobarBFRR[1,1], ArumugamBFRR[1,1], turnbaughBFRR[1,1], zupancicBFRR[1,1], WuBFRR[1,1], goodrichBFRR[1,1]), 2)
low <- signif(c(baxterBFRR[1,2], HMPBFRR[1,2], rossBFRR[1,2], escobarBFRR[1,2], ArumugamBFRR[1,2], turnbaughBFRR[1,2], zupancicBFRR[1,2], WuBFRR[1,2], goodrichBFRR[1,2]), 2)
high <- signif(c(baxterBFRR[1,3], HMPBFRR[1,3], rossBFRR[1,3], escobarBFRR[1,3], ArumugamBFRR[1,3], turnbaughBFRR[1,3], zupancicBFRR[1,3], WuBFRR[1,3], goodrichBFRR[1,3]), 2)

@@

**Figure 3: Meta Analysis of the Relative Risk of Obesity Based on Bacteroidetes/Firmicutes Ratio.**  Groups were divided for each study on high and low B/F ratio groups based on the median for that specific study.  The overall pooled relative risk was 0.86 for the low diversity group (P-value = 0.0376).

@@

#Load in BMI data
data1 <- read.csv("combined.H.corr.data.csv")
#Load in Turnbaugh data
turn.data <- read.csv("Turnbaugh.normalized.csv")

#Turnbaugh STD = 0.4818751
turn.ZScore.H <- turn.data$H.corr/0.4818751

#Amish STD = 0.5932651
keepamish <- which(data1$dataset == "Amish")
amishdata <- data1[keepamish, ]
amish.ZScore.H <- amishdata$H.corr/0.5932651

#MetaHit-Den STD = 0.4038558
keepmetahit <- which(data1$dataset == "MetaHit")
metahitdata <- data1[keepmetahit, ]
metahit.ZScore.H <- metahitdata$H.corr/0.4038558

#Korean STD = 0.5400615
keepKorean <- which(data1$dataset == "Korean")
Koreandata <- data1[keepKorean, ]
Korean.ZScore.H <- Koreandata$H.corr/0.5400615

#HMP STD = 0.6774514
keepHMP <- which(data1$dataset == "HMP")
hmpdata <- data1[keepHMP, ]
HMP.ZScore.H <- hmpdata$H.corr/0.6774514

#CCHC STD = 0.5784522
keepCCHC <- which(data1$dataset == "CCHC")
CCHCdata <- data1[keepCCHC, ]
CCHC.ZScore.H <- CCHCdata$H.corr/0.5784522

#GLNE07 STD = 0.4663871
keepGLNE <- which(data1$dataset == "GLNE")
GLNEdata <- data1[keepGLNE, ]
GLNE.ZScore.H <- GLNEdata$H.corr/0.4663871

#Escobar STD = 0.5378935
keepescobar <- which(data1$dataset == "Escobar")
escobardata <- data1[keepescobar, ]
Escobar.ZScore.H <- escobardata$H.corr/0.5378935

#COMBO STD = 0.6870763
keepCOMBO <- which(data1$dataset == "COMBO")
COMBOdata <- data1[keepCOMBO, ]
COMBO.ZScore.H <- COMBOdata$H.corr/0.6870763

#COMBO STD = 0.3489582
keeptwinsUK <- which(data1$dataset == "twinsUK")
twinsUKdata <- data1[keeptwinsUK, ]
twinsUK.ZScore.H <- twinsUKdata$H.corr/0.3489582

#data1 order Escobar, GLNE, CCHC, HMP, Korean, MetaHit, COMBO, twinsUK
HZScore <- c(Escobar.ZScore.H, GLNE.ZScore.H, CCHC.ZScore.H, HMP.ZScore.H, Korean.ZScore.H, metahit.ZScore.H, COMBO.ZScore.H, twinsUK.ZScore.H, amish.ZScore.H)
data1 <- cbind(HZScore, data1)

turn.data <- cbind(turn.ZScore.H, turn.data)
rm(Escobar.ZScore.H, GLNE.ZScore.H, CCHC.ZScore.H, HMP.ZScore.H, Korean.ZScore.H, metahit.ZScore.H, amish.ZScore.H, turn.ZScore.H, COMBO.ZScore.H, twinsUK.ZScore.H)
rm(keepmetahit, keepCCHC, keepescobar, keepGLNE, keepHMP, keepKorean)

@@

#Create group analysis table to harness full data set
#For ZScore normalization
d1.obese <- data1$obese
t.obese <- as.character(turn.data$obese)

@@

d1.H.Zscore <- data1$HZScore
t.H.Zscore <- turn.data$turn.ZScore.H

#Combine everything together
comb1.obese <- c(d1.obese, t.obese)
comb1.bmi.class <- c(d1.bmi.class, t.bmi.class)
comb1.H.Zscore <- as.data.frame(c(d1.H.Zscore, t.H.Zscore))

group.data.test1 <- cbind(comb1.H.Zscore, comb1.obese, comb1.bmi.class)
colnames(group.data.test1) <- c("H.Zscore", "obese", "BMI.class")

@@

##Run Zscore normalized Bacter and Firm overall analysis
#data1 order Escobar, GLNE, CCHC, HMP, Korean, MetaHit
d1.zbf.corr <- data2$ZBF
t.zbf.corr <- turn.data$ZBF

@@

@@

```{r Figure_5, echo=FALSE, message=FALSE, warning=FALSE}
# Section A
p1 <- ggplot(group.data.test, aes(obese, H.Zscore)) 
p2 <- p1 + geom_boxplot(aes(fill = obese)) + scale_fill_brewer(palette="Set1")
p3 <- p2 + theme(axis.text=element_text(size=8), axis.title=element_text(size=10, face="bold"))
p4 <- p3 + xlab("Obese Category") + ylab("ZScore Normalized H") 
p5 <- p4 + guides(fill=guide_legend(title="Obese")) + theme(legend.title = element_text(size=8, face="bold"))
p6 <- p5 + theme(axis.line=element_line(colour="black"), 
                 panel.grid.major = element_blank(), 
                 panel.grid.minor = element_blank(), 
                 panel.border = element_blank(), 
                 panel.background = element_blank())
p7 <- p6 + ggtitle('A') + theme(plot.title=element_text(hjust=0, face="bold", size=20))
